### 2020-01-12 ForReport20200114/HandOutEngBayes.Rmd

---
title: "Try Bayesian Survival Analysis using the JACC study data (Bayesian Survival Analysis presentations)"
subtitle: "Tables and figures for analysing the association between milk intake and stroke mortality."
author: "Chaochen Wang | Aichi Medical University"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
    keep_tex: true
    # includes:  
    #   in_header: preamble-latex.tex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

## Parametric Survival Analysis Model (in a Bayesian way)

```{marginfigure}
Weibull Survival funciton in a proportional hazard way expression is:
$$S(t | \mathbf{x}) = \exp\{ -\lambda t ^\kappa e^{\mathbf{\beta_{PH}^T x}} \}$$
```

```{marginfigure}
Weibull Survival funciton in a AFT way expression is:
$$S(t | \mathbf{x}) = \exp\{ -\lambda t ^\kappa e^{\kappa\mathbf{\beta_{AFT}^T x}} \}$$
```

- We chose Weibull model to describe the survival data we have.

- Weibull model is both a proportional hazard model and a accelerated failure time (AFT) model. The relationship between the acceleration factor and log hazard ratio:

$$
\exp(\beta^T_{PH}) = \exp(\kappa\beta^T_{AFT})
$$

## An example of a crude Weibull model using the JAGS program in R:

```{r modeljag, eval=FALSE}
jacc.weibull.model <- function() {
  for(i in 1:39386){
    # sAge[i] <- (Age[i] - mean(Age[])) / sd(Age[])
    is.censored[i] ~ dinterval(t[i], c[i])
    t[i] ~ dweib(shape, lambda[i])
    lambda[i] <- exp(-mu[i] * shape)
    mu[i] <- beta[1] + beta[2]*equals(Mlkfre[i], 2) +
      beta[3]*equals(Mlkfre[i], 3) + beta[4] * equals(Mlkfre[i], 4) +
      beta[5] * equals(Mlkfre[i], 5) # + 
      # beta[6] * sAge[i]

    ### calculate log-likelihoods
    # y[i] <- ifelse(is.censored[i], c[i], t[i])
    # loglik[i] <- log(ifelse(is.censored[i],
    #                  exp(-lambda[i] * (y[i] ^ shape)),
    #               shape * lambda[i] * (y[i] ^ (shape - 1)) * exp(-lambda[i] * (y[i] ^ shape))))
  }

  ## priors for betas
  for(j in 1:5){
    beta[j] ~ dnorm(0, 0.001)
  }

  ### prior for shape
  shape ~ dgamma(.001, .001)

  ### Generated values
  AFT[2] <- exp(beta[2])
  HR[2] <- exp(shape * beta[2])
  p.crit[2] <- step(1 - HR[2])

  AFT[3] <- exp(beta[3])
  HR[3] <- exp(shape * beta[3])
  p.crit[3] <- step(1 - HR[3])

  AFT[4] <- exp(beta[4])
  HR[4] <- exp(shape * beta[4])
  p.crit[4] <- step(1 - HR[4])

  AFT[5] <- exp(beta[5])
  HR[5] <- exp(shape * beta[5])
  p.crit[5] <- step(1 - HR[5])
}
```

You can use the following codes to run the above model (careful about the cost of time):

```{r eval=FALSE}
library(R2jags) #conect to jags sampling engine in R
jagsfit <- jags.parallel(data = JACCdata,
                         parameters.to.save = c("beta[5]", "HR[5]", "AFT[5]"),
                         #add more variables if you want to check more
                         n.iter = 10000, 
                         n.burnin = 2500, 
                         n.chains = 3,
                         model.file = jacc.weibull.model)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("../data/jagmodel0men.Rdata")
# Simulated <- coda::as.mcmc(jagsfit)
Simulated <- mcmcplots::as.mcmc.bugs(jagsfit$BUGSoutput)
library(ggmcmc)
ggSample <- ggs(Simulated)

```

## Check model fitting:

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.margin = FALSE}
library(tidyverse)
ggSample %>% 
  filter(Iteration >= 500 & Parameter %in% c("beta[5]", "HR[5]", "AFT[5]")) %>% 
  ggs_traceplot()
```